Ornaments as indicators of Social change before and after European contact at Kiwulan, Northeastern Taiwan

### 2019-06-22 analysis/paper/paper.Rmd
code for nndist hypothesis testing and similation, make repetitive code shorter by map()function

```{r read-in-the-spatial-data, eval = TRUE}
library(sf)
library(cartography)
burial <- st_read(here("analysis", "data", "raw_data", "AD_burial.shp"))
AD <- st_read(here("analysis", "data", "raw_data", "AD_zone.shp"))
AD_data <- st_read(here("analysis", "data", "raw_data", "AD_withdata.shp"))
location <- st_read(here("analysis", "data", "raw_data", "location.shp"))
units_40 <- read.csv(here("analysis", "data", "raw_data", "kwl-list-of-sampling-squares.csv"))
```

@@

```{r point-pattern-data-setup}
# get centroid 
ornaments_period_join_cen <-
  ornaments_period_join %>% 
  mutate(lon = map_dbl(geometry, ~st_point_on_surface(.x)[[1]]),
         lat = map_dbl(geometry, ~st_point_on_surface(.x)[[2]]))

# plot the centroid
ggplot() +
  geom_sf(data = sample_units, fill = NA) +
  geom_point(data = ornaments_period_join_cen, 
             aes(x = lon, y = lat))

# transform data to PPP format and sp SpatialPoints
library(spatstat)

#store x and y coords in two vectors, and create two vectors xrange and yrange with dimensions of triangle that contain all points
lon <- ornaments_period_join_cen$lon
lat <- ornaments_period_join_cen$lat
xrange <- range(lon, na.rm=T)
yrange <- range(lat, na.rm=T)

# create ppp
ornaments_point <- ppp(lon, lat, xrange, yrange)

To access whether the distribution of ornaments present different density that can indicate the changes caused by global trade network brought by the Europeans, point pattern analysis is used to measure and test the density of ornaments across three time periods.

>>>>

```{r read-in-the-spatial-data, eval = TRUE}
library(sf)
library(cartography)
burial <- invisible(st_read(here("analysis", "data", "raw_data", "AD_burial.shp"), quiet = TRUE))
AD <- invisible(st_read(here("analysis", "data", "raw_data", "AD_zone.shp"), quiet = TRUE))
AD_data <- invisible(st_read(here("analysis", "data", "raw_data", "AD_withdata.shp"), quiet = TRUE))
location <- invisible(st_read(here("analysis", "data", "raw_data", "location.shp"), quiet = TRUE))
units_40 <- invisible(read.csv(here("analysis", "data", "raw_data", "kwl-list-of-sampling-squares.csv")))
```

@@

```{r point-pattern-data-setup}
# get centroid and make a list of three periods
ornaments_period_join_cen_all <-
  ornaments_period_join %>% 
  mutate(lon = jitter(map_dbl(geometry, ~st_point_on_surface(.x)[[1]])),
         lat = jitter(map_dbl(geometry, ~st_point_on_surface(.x)[[2]])))

ornaments_period_join_cen <- 
 ornaments_period_join_cen_all %>% 
  group_by(period) %>% 
  group_split() %>% 
  set_names(c("Pre-European",
              "European",
              "Chinese"))
```

```{r}
# transform data to PPP format and sp SpatialPoints
library(spatstat)
library(maptools)
library(raster)

#store x and y coords in two vectors, and create two vectors xrange and yrange with dimensions of triangle that contain all points, by period

# use common range for each period
lon_all <- ornaments_period_join_cen_all$lon
lat_all <- ornaments_period_join_cen_all$lat
xrange <- range(lon_all, na.rm=T)
yrange <- range(lat_all, na.rm=T)

# create a function to apply to each period
density_analysis <- function(x, ...) {
  # create ppp and sp(spacial points) format
lon_tmp <- x$lon
lat_tmp <- x$lat
ornaments_point_ppp_tmp <- ppp(lon_tmp, lat_tmp, xrange, yrange)
ornaments_point_sp_tmp <- as(ornaments_point_ppp_tmp, "SpatialPoints")

# kernel density estimation with contour
kernel_tmp <- density(ornaments_point_ppp_tmp)
kernel_to_plot <- 
  plot(kernel_tmp, ...)
  contour(kernel_tmp, add = TRUE)

return(list(ornaments_point_ppp = ornaments_point_ppp_tmp,
            kernel_to_plot = kernel_to_plot))
}
```

```{r plot-all-densities}
par(mfrow = c(1, 3))
# apply function to each data frame in the list
titles <- names(ornaments_period_join_cen)

kernels_for_plotting <- 
  map2(ornaments_period_join_cen, 
       titles,
      ~density_analysis(.x, main = .y))
```

```{r}
# function for hypothesis testing
nndist_hypo_testing <- function(x, n = 1e3L,  ... ) {

ann_tmp <- mean(nndist(x$ornaments_point_ppp, k = 1))

# object for storing simulated ANN values
ann_r_tmp <- vector(length = n)

for (i in 1:n) {
  # Generate random point locations
  rand_tmp <-  rpoint(n = x$ornaments_point_ppp$n,
                      win = as.owin(as_Spatial(sample_units)))
  ann_r_tmp[i] <-
    mean(nndist(rand_tmp, k = 1))  # Tally the ANN values
}

the_plot <- 
hist(
  ann_r_tmp,
  xlab = NULL,
  las = 1,
  breaks = 40,
  col = "bisque",
  xlim = range(ann_tmp, ann_r_tmp),
  ...)

abline(v = ann_tmp, col = "blue")

return(the_plot)
}
```

```{r}

par(mfrow = c(3, 1))
invisible(
map2(kernels_for_plotting, 
    titles,
    ~invisible(nndist_hypo_testing(.x, 
                         main = .y)))
)
```

To assess whether the distribution of ornaments present different density that can indicate the changes caused by global trade network brought by the Europeans, point pattern analysis is used to measure and test the density of ornaments across three time periods. We assume a uniform point density across the research area before European contact period, that is, randomly distributed, and test the null hypothesis that is there is no cluster of ornaments across the sapce during the European contact periods.
